Balance responses in MS

### 2020-02-04 R/msBalance.Rmd
changed name of norm coef tibble from taRNorm to taRMax

figureNumbers <- captioner(prefix = "Figure")
captionCartoon = figureNumbers(name = "cartoon", caption = "Perturbation directions used in testing protocol.")
captionRawEmg = figureNumbers(name = "rawEmg", caption = "EMG traces stratified by perturbation direction and study group, prior to normalization. Colors represent individual participants.")
```

@@

# to calculate normalization coefficients, average across like trials for each patient, limit to a particular normalization window, and calculate maximum values
taRNorm = taRAv %>% filter(Time %between% c(0.0, 1.0)) %>% group_by(Group, Participant, `Perturbation Direction`) %>% summarize(Vm = max(V, na.rm = TRUE)) %>% ungroup()
```

***
```{r rawEMG, out.width = "100%", fig.align = "left", fig.cap = captionRawEmg}
p = ggplot(taR, aes(x = Time, y = V)) + geom_line(aes(color = Participant), show.legend = FALSE, alpha = 0.5) + facet_grid(Group ~ `Perturbation Direction`) + theme_minimal_grid() +
  geom_jitter(mapping = aes(x = 0, y = Vm, color = Participant), data = taRNorm %>% left_join(participants %>% select(Participant, Group)), show.legend = FALSE, height = 0, width = 0.01)
p
```

```{r emgMags}
emgMags = 
  taRNorm %>% group_by(`Perturbation Direction`, Group) %>% summarize(Mean = mean(Vm) %>% round() ) %>% left_join(
    taRNorm %>% group_by(`Perturbation Direction`, Group) %>% summarize(SD = sd(Vm) %>% round() ) ) %>% left_join(
      taRNorm %>% group_by(`Perturbation Direction`, Group) %>% summarize(Max = max(Vm) %>% round() ) )
emgMags %>% kable(caption = captionEmgMags) %>% kable_styling(full_width = F, position = "left")

# 
# 
# ggplot(taRAv, aes(Time, Vn, color = Participant)) + geom_point(show.legend = FALSE) + facet_wrap(vars(`Perturbation Direction`))
# 
# p = ggplot(taRAv, aes(Time, Vn)) + geom_point()
# 
# p = ggplot(taRAv %>% group_by(Participant, `Perturbation Direction`), aes(x = Time, y = Vn, color = Participant)) +
#   facet_grid(Group ~ `Perturbation Direction`) +
#   theme_minimal_grid() +
#   coord_cartesian(xlim = c(0,1)) + 
#   geom_line(aes(group = Participant, color = Participant), taRav, show.legend = FALSE) +
# p
```

***

>>>>

figureNumbers <- captioner(prefix = "Figure")
captionCartoon = figureNumbers(name = "cartoon", caption = "Perturbation directions used in testing protocol.")
captionRawEmg = figureNumbers(name = "rawEmg", caption = "EMG traces stratified by perturbation direction and study group, prior to normalization. Gray traces represent individual trials. Colors represent averages of individual participants.")
captionNormedEmg = figureNumbers(name = "normedEmg", caption = "EMG traces stratified by perturbation direction and study group, after normalization. Colors represent averages for individual participants.")
```

@@

# to calculate normalization coefficients, average across like trials for each patient, limit to a particular normalization window, and calculate maximum values
taRMax = taRAv %>% filter(Time %between% c(0.0, 1.0)) %>% group_by(Group, Participant, `Perturbation Direction`) %>% summarize(Vm = max(V, na.rm = TRUE)) %>% ungroup()
```

***
```{r rawEMG, out.width = "75%", fig.align = "left", fig.cap = captionRawEmg}
p = ggplot(taR, aes(x = Time, y = V)) +
  geom_line(color = "Gray90", show.legend = FALSE, alpha = 0.5) +
  geom_line(aes(color = Participant), taRAv, show.legend = FALSE, alpha = 0.5) +
  facet_grid(Group ~ `Perturbation Direction`) +
  theme_minimal_grid() +
  geom_text(mapping = aes(x = 0, y = Vm, color = Participant, label = Participant), data = taRMax %>% group_by(Group, `Perturbation Direction`) %>% top_n(3), show.legend = FALSE)
p
```

By visual inspection, there were some gross magnitude differences in recorded EMG levels between the MS and control groups, with higher EMG levels overall in the MS group.
EMG levels in most control participants (with the exception of MSC15 and MSC18) were generally roughly an order of magnitude lower than levels in MS participants.
This number is difficult to interpret, as it can reflect a large number of biological and methodological factors, so that increased absolute level among MS could reflect increased muscle activation or completely unrelated factors.
EMG levels were therefore normalized to the maximum values obtained in a 1 second window after perturbation onset after averaging across like trials (`r tableNumbers("emgMags", display = "cite")`).

```{r emgMags}
emgMags = 
  taRMax %>% group_by(`Perturbation Direction`, Group) %>% summarize(Mean = mean(Vm) %>% round() ) %>% left_join(
    taRMax %>% group_by(`Perturbation Direction`, Group) %>% summarize(SD = sd(Vm) %>% round() ) ) %>% left_join(
      taRMax %>% group_by(`Perturbation Direction`, Group) %>% summarize(Max = max(Vm) %>% round() ) )
emgMags %>% kable(caption = captionEmgMags) %>% kable_styling(full_width = F, position = "left")
```

```{r normedEMG, out.width = "75%", fig.align = "left", fig.cap = captionNormedEmg}
p = ggplot(taR, aes(x = Time, y = V)) +
  geom_line(color = "Gray90", show.legend = FALSE, alpha = 0.5) +
  geom_line(aes(color = Participant), taRAv, show.legend = FALSE, alpha = 0.5) +
  facet_grid(Group ~ `Perturbation Direction`) +
  theme_minimal_grid() +
  geom_text(mapping = aes(x = 0, y = Vm, color = Participant, label = Participant), data = taRMax %>% group_by(Group, `Perturbation Direction`) %>% top_n(3), show.legend = FALSE)
p
```
